A simulation-based assessment of the relation between Stone Age sites and relative sea-level change along the Norwegian Skagerrak coast

### 2022-02-12 analysis/paper/paper.Rmd
Re-running analyses after IDW

The close association between Scandinavian Stone Age settlements and prehistoric shorelines was becoming apparent by the end of the 19th century [@degeer1896], and was firmly established and first applied as a dating method in Sweden [@hollender1901] and Norway [@brogger1905] at the turn of the century. Shoreline dating has been fundamental to Stone Age research in Norway ever since [see for example @berg-hansen2009; @bergsvik2020; @bjerck2008; @damlien2021]. The method is used both independently, and to compliment other sources of temporal data such as typological indicators or radiometric dates. However, given the coarse and fuzzy resolution of established typological frameworks, the vast amount of surveyed sites that only contain generic lithic debitage that could hail from any part of the period, and as the conditions for the preservation of organic material is typically poor in Norway, dating with reference to shoreline displacement is often the only and most accurate method by which one can hope to date the sites. Shoreline dating is consequently fundamental to our understanding of the Norwegian Stone Age. This is both because it is central to the temporal framework on which our understanding of the period is based, but also because the method is only applicable so long as the societies in question have continuously settled on or close to the contemporary shoreline. Consequently, adherence or deviation from this pattern also has major implications for the socio-economic foundations of the societies in question.

>>>>

The close association between Scandinavian Stone Age settlements and shifting prehistoric shorelines was becoming apparent by the end of the 19th century [@degeer1896], and was firmly established and being applied as a dating method in Sweden [@hollender1901] and Norway [@brogger1905] at the turn of the century. Shoreline dating has been fundamental to Stone Age research in Norway ever since [see for example @berg-hansen2009; @bergsvik2020; @bjerck2008; @damlien2021]. The method is used both independently, and to compliment other sources of temporal data such as typological indicators or radiometric dates. However, given the coarse and fuzzy resolution of established typological frameworks, the vast amount of surveyed sites that only contain generic lithic debitage that could hail from any part of the period, and as the conditions for the preservation of organic material is typically poor in Norway, dating with reference to shoreline displacement is often the only and most accurate method by which one can hope to date the sites. Shoreline dating is consequently fundamental to our understanding of the Norwegian Stone Age. This is both because it is central to the temporal framework on which our understanding of the period is based, but also because the method is only applicable so long as the societies in question have continuously settled on or close to the contemporary shoreline. Consequently, adherence or deviation from this pattern also has major implications for the socio-economic foundations of the societies in question.
